
The following lag determination code was provided by Associate Professor Erik Vanhatalo to whom I owe my thanks.
The code was updated with regards to some packages, but the functions are unchanged.


This code snippet contains the necessary packages to carry out lag determination via the methods developed by
Vanhatalo et al., Rato & Reis and and Ku et al. 
```{r}
library(MASS)
library(paran)
library(MTS)
library(ppcor)
library(readxl)
library(corrplot)
library(qcc)
library(forecast)
library(rARPACK)
library(elasticnet)
library(plot3D)
library(Hmisc)

par(mar=c(1,1,1,1))

###################################################
###################################################
### THE CODE STARTS WITH FUNCTIONS NEEDED FOR   ###
### THE EXAMPLES BELOW                          ###
###################################################
###################################################
####################################################### FUNCTION
################################################################
###################################################
### FUNCTION - SIMULATING VAR(1) DATA           ###
###################################################
### INPUTS:                                     ###
### p = number of variables                     ###
### obs = number of observations                ###
### burn = Number of initial observations to    ###
###        'burn'                               ###
### mu = Mean of the errors                     ###
### sigma = covariance matrix of the errors     ###
### phi = Matrix with autocorr. coeff.          ###
###                                             ###
### OUTPUT:                                     ###
###   A matrix with VAR(1) data                 ###
###################################################
var1.sim <- function(p,obs,burn,mu,sigma,phi){
  x1 = matrix(rep(0),(obs+burn),p)
  e1 = mvrnorm((obs+burn),mu,sigma)
  x1[1,] = mu
  
  for (k in 2:(obs+burn)){
    x1[k,] = phi %*% x1[(k-1),] + e1[k,] # VAR(1)
  }
  x1 = x1[(burn+1):(obs+burn),] #burns first obs.
  x1
}

####################################################### FUNCTION
################################################################
###################################################
###################################################
###           SIMULATED VMA DATA                ###
###       Using Tsay's 'MTS' package            ###
###                                             ###
### INPUTS EXPLAINED:                           ###
###   noobs = Sample size                       ###
###   arlags = The exact lags of the VAR matrix ###
###   polynomial.                               ###
###   malags = The exact lags of the VMA matrix ###
###   polynomial.                               ###
###   cnst = Constant vector, Phi0              ###
###   phi = Array of VAR coefficient matrices   ###
###   in the order of the given arlags.         ###
###   theta = Array of VMA coefficient matrices ###
###   in the order of the given malags.         ###
###   skip = The number of initial data to be   ###
###   omitted. Deafult is skip = 200.           ###
###   sigma = Covariance matrix (k-by-k,        ###
###   positive definite) of the innovations.    ###
###################################################
###################################################


####################################################### FUNCTION
################################################################
###################################################
### ADDING A SHIFT TO VAR(1) DATA IN PHASE II   ###
### THE FUNCTION CREATES BOTH PHASE I AND       ###
### PHASE II DATA                               ###
###################################################
### INPUTS:                                     ###
### p = number of variables                     ###
### mu1 = Mean of the errors (in Phase I)        ###
### sigma = covariance matrix of the errors     ###
### phi = Matrix with autocorr. coeff.          ###
### obs1 = number of obs. Phase I               ###
### burn = Number of initial observations to    ###
###        'burn'                               ###
### obs2 = number of obs. Phase II              ###
### shf = Row vector, shifts in std. dev. units ###
### shf_obs = shift intruced at a specific obs 
###           in Phase II.
### OUTPUT:                                     ###
###   A matrix with Phase I and Phase II data   ###
###################################################
VAR1_shift_PI_PII <-function(p,mu,sigma,phi,obs1,burn,obs2,shf,shf_obs){
  
  ecov<-solve(diag(p^2)-kronecker(phi, phi)) # Inv. of expression in Reinsel (1993)
  stsigma<-cbind(stack(as.data.frame(sigma))[,1]) # Stacking sigma
  dfg<-ecov%*%stsigma
  ncov<-matrix(c(dfg),p,p) # Creating matrix again, the "true" covariance matrix.
  mu2 <- as.vector(mu1+shf*sqrt(diag(ncov))) # New mean of the process.
  errshf = as.vector((diag(p)-phi)%*%mu2)
  
  # Simulating data in Phase I
  
  x1 = matrix(rep(0),(obs1+burn),p)
  e1 = mvrnorm((obs1+burn),mu,sigma)
  x1[1,] = mu
  
  for (k in 2:(obs1+burn)){
    x1[k,] = phi %*% x1[(k-1),] + e1[k,] # VAR(1)
  }
  x1 = x1[(burn+1):(obs1+burn),] #burns first obs.
  
  
  # Simulating data in Phase II
  x2 = matrix(rep(0),obs2,p)
  e21 = mvrnorm((shf_obs-1),mu,sigma) # Errors before shift in Phase II
  e22 = mvrnorm(obs2,errshf,sigma) # Errrors after shift
  
  if (shf_obs < 2) {
    x2[1,] = phi %*% x1[obs1,] + e22[1,] # First value in Phase II
    
    for (k in 2:obs2){
      x2[k,] = phi %*% x2[(k-1),] + e22[k,]
    }
    
  }
  
  if (shf_obs > 1) {
    x2[1,] = phi %*% x1[obs1,] + e21[1,] # First value in Phase II
    
    for (k in 2:(shf_obs-1)){
      x2[k,] = phi %*% x2[(k-1),] + e21[k,] # Shift has not yet occured
    }
    for (k in shf_obs:obs2){
      x2[k,] = phi %*% x2[(k-1),] + e22[k,] # Shift has occurred
    }
  }
  
  res = matrix(rep(0),(obs1+obs2),p) #Combining Phase I and Phase II data
  res[1:obs1,]=x1
  res[(obs1+1):(obs1+obs2),]=x2
  res
}


####################################################### FUNCTION
################################################################
###################################################
### ADDING A SHIFT TO VMA(1) DATA IN PHASE II   ###
### THE FUNCTION CREATES BOTH PHASE I AND       ###
### PHASE II DATA                               ###
###################################################
### INPUTS:                                     ###
### p = number of variables                     ###
### mu1 = Mean of the errors (in Phase I)       ###
### sigma = covariance matrix of the errors     ###
### theta = Matrix with MA(1) coeff.            ###
### obs1 = number of obs. Phase I               ###
### burn = Number of initial observations to    ###
###        'burn'                               ###
### obs2 = number of obs. Phase II              ###
### shf = Row vector, shifts in std. dev. units
### shf_obs = shift intruced at a specific obs 
###           in Phase II.
### OUTPUT:                                     ###
###   A matrix with Phase I and Phase II data   ###
###################################################
VMA1_shift_PI_PII <-function(p,mu1,sigma,theta,obs1,burn,obs2,shf,shf_obs) {
  #Calculating shift size in terms of true standard deviations of the variables
  
  covMA1 = sigma +(theta %*% sigma %*% t(theta)) #true Covariance matrix
  mu2 <- as.vector(mu1+(shf*sqrt(diag(covMA1)))) #new mean of the process
  ##errshf = mu2as.vector(mu2 %*% (diag(p)%/%theta)) 
  
  # Simulating data in Phase I
  
  x1 = matrix(rep(0),(obs1+burn),p)
  e1 = mvrnorm((obs1+burn),mu1,sigma)
  
  x1[1,] = mu1 +e1[1,]
  
  for (k in 2:(obs1+burn)){
    x1[k,] = mu1 + e1[k,] - theta %*% e1[(k-1),]
  }
  x1 = x1[(burn+1):(obs1+burn),] #burns first obs.
  
  # Simulating data in Phase II
  x2 = matrix(rep(0),obs2,p)
  e2 = mvrnorm(obs2,mu1,sigma) # New errors in Phase II (shifts in mean of errors)
  
  if (shf_obs < 2) {
    x2[1,] = mu2 + e2[1,]- theta %*% e1[obs1,] # First value in Phase II (connecting with Phase I)
    
    for (k in 2:obs2){
      x2[k,] = mu2 + e2[k,] - theta %*% e2[(k-1),]
    }
  }
  
  if (shf_obs > 1) {
    x2[1,] = mu1 + e2[1,]- theta %*% e1[obs1,] # First value in Phase II (connecting with Phase I)
    
    for (k in 2:(shf_obs - 1)) {
      x2[k,] = mu1 + e2[k,] - theta %*% e2[(k-1),]
    }
    for (k in shf_obs:obs2){
      x2[k,] = mu2 + e2[k,] - theta %*% e2[(k-1),]
    }
  }
  
  
  res = matrix(rep(0),(obs1+obs2),p) #Combining Phase I and Phase II data
  res[1:obs1,]=x1
  res[(obs1+1):(obs1+obs2),]=x2
  res
}

####################################################### FUNCTION
################################################################
###################################################
### FUNCTION - ADDING LAGS TO A MATRIX          ###
###################################################
### INPUTS:                                     ###
###   X = Original matrix                       ###
###   lags = Number of lags to add              ###
### OUTPUT:                                     ###
###   A lagged matrix, 'trimmed' due to lagging ###
###################################################
Xlag <- function(X,lags){
  p = dim(X)[2] # Number of variables.
  obs = dim(X)[1] # Number of obs.
  
  for (i in 1:lags){
    
    Xlag = matrix(rep(0),obs+i,(p*i)+p)
    
    for (l in 0:i){
      Xlag[(l+1):(obs+l),((l*p)+1):((l*p)+p)]=X
    }
    
    Xlag = Xlag[(i+1):obs,] #trims the matrix
  }
  Xlag
}


####################################################### FUNCTION
################################################################
###################################################
### PERCENT OF VARIANCE EXPLAINED               ###
### BY PCA (CORRELATION MATRIX                  ###
###################################################
### INPUTS:                                     ###
### X = Data Matrix                             ###
###                                             ###
### OUTPUT:                                     ###
###   A plot with the cumulative variance       ###
###   explained of the PCs                      ###
###################################################
pca.var.expl <- function (X){
  
  eigval = (princomp(X, cor=TRUE)$sdev)^2
  
  Sum_eigval = sum(eigval)
  
  percent = (eigval/Sum_eigval)*100
  
  Sum_percent = rep(0,length(eigval))
  Sum_percent[1]=percent[1]
  
  for (i in (2:(length(eigval)))){
    Sum_percent[i]=Sum_percent[i-1] + percent[i]  
  }
  #plot(Sum_percent,type="o",xlab="Eigenvalue number",
  #    ylab = "% variance explained")
  #title(main="Variance explained")
  Sum_percent
}


####################################################### FUNCTION
################################################################
###################################################
### FUNCTION - KU ET AL. (1995) LAG METHOD      ###
###################################################
### INPUTS:                                     ###
### X = Original data matrix                    ###
### The parallel analysis inside the function   ###
### uses the 'paran' package and its default    ###
### values                                      ###
###                                             ###
### OUTPUT:                                     ###
###   A summarizing table with the information  ###
###   in the steps of the procedure and a       ###
###   numeric value with the lags in the        ###
###   numeric part of the procedure             ###
###################################################

ku.et.al.1995 <- function(X){
  
  p = dim(X)[2] # Number of variables in X
  save_r_l = rep(0,dim(X)[1])
  save_r_new_l = rep(0,dim(X)[1])
  
  summary = matrix(rep(0),10 ,5)
  colnames(summary) = c("lags (l)","Variables", "PCs retained",
                        "r(l)","r_new(l)")
  # 1. Set l = 0
  lags = 0
  
  iter = 1 # Count variable to run/stop while loop.
  
  while(iter>0){
    
    # 2. Form data matrix
    x <- X
    
    if(lags>0){
      x <-Xlag(X,lags)
    }
    
    # 3. Perform PCA and calculate scores
    #    This step is not needed when using 'paran' package in Step 4.
    
    # 4. Set j and r (number of linear relations)
    j = p*(lags+1)
    r_l = 0
    
    # 5.Determine if the j:th component is a linear relation 
    #   "r_l" is determined such that r_l = number of "low" eigenvalues - 1
    #   in each stage. See p.186 in Ku et al. (1995).
    #   The number of "low" eigenvalues can be determined by e.g. parallell analysis; 
    #   here using the 'paran' package. 
    
    par = paran(x,quietly=TRUE, status=FALSE)$Retained
    
    r_l = (dim(x)[2]-par)
    
    save_r_l[lags+1]=r_l
    
    # 7. Number of new relationships r_new_l
    if (lags==0){
      r_new_l = r_l
      save_r_new_l[1]=r_new_l
      summary[1,1]=0
      summary[1,2]=dim(x)[2]
      summary[1,3]=par
      summary[1,4:5]=r_l
    }
    
    if (lags==1){
      r_new_l = (r_l - 2*save_r_new_l[1])
      save_r_new_l[2]=r_new_l
      summary[2,1]=1
      summary[2,2]=dim(x)[2]
      summary[2,3]=par
      summary[2,4]=r_l
      summary[2,5]=r_new_l
    }
    
    if (lags>1){
      
      sum = c(0)
      for(i in 0:(lags-1)){
        sum = sum + (lags-i+1)*save_r_new_l[i+1]
      }  
      
      r_new_l = r_l-sum
      save_r_new_l[lags+1]=r_new_l
      
      summary[(lags+1),1]=lags
      summary[(lags+1),2]=dim(x)[2]
      summary[(lags+1),3]=par
      summary[(lags+1),4]=r_l
      summary[(lags+1),5]=r_new_l
    }
    
    # 8. If r_new(l) is less or equal to zero, go to 10. else proceed.
    if (r_new_l <= 0){
      iter = 0
      lags = lags-1
      break
    }
    
    # 9. Update lags and go to step 2
    lags = lags+1
    
    # 10. Stop
  } # Closes while loop!  
  
  # Calculate PC scores based on chosen lag + 1
  x = Xlag(X,lags+1)
  PCA = princomp(x,cor=TRUE)$scores
  
  #Creating a cross correlation funtion plot among scores
  #par(mfrow = c((lags+2),1))
  #for (i in 1:(lags+2)){
  #  ccf(PCA[,1],PCA[,(i*dim(X)[2])])  
  #}
  
  print(summary[1:(lags+2),])
  lags
}



####################################################### FUNCTION
################################################################
###################################################
### FUNCTION - RATO & REIS (2013) LAG METHOD    ###
###################################################
### INPUTS:                                     ###
### X = Original data matrix                    ###
### laglimit = limit for the number of lags to  ###
###            test                             ###
###                                             ###
### OUTPUT:                                     ###
###   Plots with Key Singular values, the       ###
###   Key singular value ratios and the         ###
###   objective function. A numeric value with  ###
###   the maximum number of lags is given.      ###
###################################################
Rato.Reis.2013 <-function(X,laglimit){
  
  # Step 1. Set lag = 0 (l = 0)
  
  m = dim(X)[2] # Number of original variables.
  
  KSV = rep(0,laglimit+1) #Saving all Key Singular Values
  KSVR = rep(0,laglimit+1) #Saving Key Singular Value Ratios
  KSV[1] = eigen(cov(X))$values[1] #First Key Singular value
  #(eigenvalue of covariance matrix)
  
  for (i in 1:(laglimit)){
    
    Xl <- Xlag(X,i) # Step 2 in Pseudocode
    KSV[i+1] = eigen(cov(Xl))$values[(m*i)+1] # Step 3&4 in Pseudocode
    if(i==1){
      KSVR[i+1] = KSV[2]/KSV[1]
    }
    if(i>1){
      KSVR[i+1] = KSV[i]/KSV[i-1] #Step 5 in Pseudocode
    }
    
  }
  is.na(KSVR) <-1 #Setting first KSVR value to "NA"
  
  KSVn = rep(0,laglimit+1) #Normalized Key Singular Values
  
  for(k in 1:(laglimit+1)){
    KSVn[k]=(KSV[k]-min(KSV))/(max(KSV)-min(KSV))
  }
  
  KSVRn = rep(0,laglimit+1) #Normalized Key Singular Value Ratios
  maxKSVR =max(KSVR[2:(laglimit+1)]) # To avoid the "NA" value
  minKSVR =min(KSVR[2:(laglimit+1)])
  
  for(k in 2:(laglimit+1)){
    KSVRn[k]=(KSVR[k]-minKSVR)/(maxKSVR - minKSVR)
  }
  is.na(KSVRn) <-1 #Setting first KSVR value to "NA"
  
  
  # Step 8 in the Pseudocode
  objfunc = rep(0,laglimit+1)
  
  for(k in 2:(laglimit+1)){
    objfunc[k] = sqrt((KSVn[k]^2)+(KSVRn[k]^2))
  }
  is.na(objfunc) <-1
  
  maxlag = which.min(objfunc)-1 #Detemines which lag minimizes objective function?
  
  x=c(0:laglimit) # For the x-axis in plots
  
  # One figure with 3 rows and 1 column
  par(mfrow=c(3,1)) 
  plot(x,KSV, type="p", xlab="No. of lags", ylab="Key Singular Value, KSV")
  plot(x,KSVR, type="p", xlab="No. of lags", ylab="Key Singular Value Ratio, KSVR")
  plot(x,objfunc, type="p", xlab="No. of lags", ylab="Objective function")
  par(mfrow=c(1,1))  
  
  maxlag
}


####################################################### FUNCTION
################################################################
###################################################
### FUNCTION - EIGENVALUES OF LAGGED            ###
### AUTOCORRELATION MATRICES BASED ON TSAY'S    ###
### CODE                                        ###
###################################################
### INPUTS:                                     ###
### X = Original data matrix                    ###
### lags = Number of lags to test               ###
### max_eig = Number of eigenvalues to plot in  ### 
###           the figure                        ###
###                                             ###
### OUTPUT:                                     ###
###   A plot of the Moduli of the eigenvalues   ###
###   at different lags                         ###
###                                             ###
### WARNINGS: You will get a warning in the     ###
###           function if all eiganvalues are   ###
###           plotted since the 'rARPACK'       ### 
###           package warns and suggest 'eigen' ###
###           to be used instead.               ###
###################################################
eig.ccm <-function(X,lags,max_eig){
  p = dim(X)[2]
  
  # For independent data to add help line in plot later
  mu_ind = rep(0,p)
  obs_ind = dim(X)[1]
  sigma_ind = cor(X)
  phi_ind = diag(p)*0
  burn_ind = 500
  Z= var1.sim(p,obs_ind,burn_ind,mu_ind,sigma_ind,phi_ind)
  
  
  A<-as.array(ccm(X, lags = lags, level = TRUE, output = F)) 
  B <- as.array(ccm(Z, lags = lags, level = TRUE, output = F)) 
  
  ccm<-A$ccm # A matrix with the components of the correlation matrices 
  ccm_ind <-B$ccm
  
  eigval_k <- array(0, dim = c(lags+1,max_eig))
  eigval_k_ind <- array(0, dim = c(lags+1,max_eig))
  #corrmat_eig_k <- array(0, dim = c(lags+1,p,p))
  
  #'LA' below = "The k largest (algebraic) eigenvalues, considering any negative sign.
  
  for (i in 1:(lags+1)){
    cormat_i <- matrix(ccm[,i],p,p)
    cormat_i_ind <- matrix(ccm_ind[,i],p,p)
    eigval_i <- Mod(eigs_sym(cormat_i,max_eig, which="LA",
                             opts=list(retvec=FALSE))$values)
    #eigvec_i <- eigen(cormat_i)$vectors
    eigval_i_ind <- Mod(eigs_sym(cormat_i_ind,max_eig, which="LA",
                                 opts=list(retvec=FALSE))$values)
    eigval_k[i,] <- eigval_i
    #corrmat_eig_k[i,,] <- eigvec_i
    eigval_k_ind[i,] <- eigval_i_ind
  }
  
  #Maximum of the eigenvalues for independent data in each lag (not zero lag)
  max_eig_ind_lag = max(eigval_k_ind[2:(lags+1),])
  
  
  par(mfrow=c(1,1))
  v1 <- seq(0,lags,5) # defines position of tick marks.
  matplot(0:lags,eigval_k,type="n",,xlab="lag",ylab="AVE", axes=FALSE)
  axis(1,at = v1,tck=-.05)
  axis(2,las=2)
  minor.tick(nx = 5,ny = 1,tick.ratio=0.75)
  
  grid(col = "lightgray", lty = "dotted") # To get grid behind points('type='n')
  par(new=TRUE)
  matplot(0:lags,eigval_k,type="b",xaxt="n",yaxt="n",ann=FALSE)
  abline(h = max_eig_ind_lag, lty="dashed", lwd=2)
  
} #Closes function


####################################################### FUNCTION
################################################################
###################################################
### EIGENVALUES OF LAGGED PARTIAL               ###
### AUTOCORRELATION MATRICES                    ###
###################################################
### INPUTS:                                     ###
### X = Data Matrix                             ###
### lags = Number of lags to test               ###
### max_eig = Maximum number of eigenvalues     ###
###           in the plot                       ###
###                                             ###
### OUTPUT:                                     ###
###   A plot of the eigenvalues at different    ###
###   lags                                      ###
###                                             ###
###################################################
eig.p.corr <- function(X,lags,max_eig){
  p = dim(X)[2]
  
  # For independent data to add help line in plot later
  mu_ind = rep(0,p)
  obs_ind = dim(X)[1]
  sigma_ind = cor(X)
  phi_ind = diag(p)*0
  burn_ind = 500
  Z= var1.sim(p,obs_ind,burn_ind,mu_ind,sigma_ind,phi_ind)
  
  savelagcor <- array(0, dim = c(lags,p,p))
  saveeigpcor_i <- array(0, dim = c(lags,max_eig))
  savelagcor_ind <- array(0, dim = c(lags,p,p))
  saveeigpcor_i_ind <- array(0, dim = c(lags,max_eig))
  
  for (i in 1:lags){
    Xlag = Xlag(X,i) #Creates a lagged matrix of data
    Xlag_ind = Xlag(Z,i)
    pcor_lag = pcor(Xlag)$estimate
    pcor_lag_ind = pcor(Xlag_ind)$estimate
    
    pcormat = matrix(rep(0),p,p) #Starting a new p.corr.matrix
    pcormat_ind = matrix(rep(0),p,p)
    for (j in 1:p){
      for (k in 1:p){
        pcormat[j,k] =  pcor_lag[j,((i*p)+k)]
        pcormat_ind[j,k] =  pcor_lag_ind[j,((i*p)+k)]# j = row, k = column
      }   
    } 
    savelagcor[i,,] = pcormat
    savelagcor_ind[i,,] = pcormat_ind
    saveeigpcor_i[i,] = Mod(eigs(pcormat,max_eig, which="LR",
                                 opts=list(retvec=FALSE))$values)
    saveeigpcor_i_ind[i,] = Mod(eigs(pcormat_ind,max_eig, which="LR",
                                     opts=list(retvec=FALSE))$values)
  }
  
  #Maximum of the eigenvalues for independent data in each lag
  max_eig_ind_lag = max(saveeigpcor_i_ind[2:(lags),])
  
  par(mfrow=c(1,1))
  v1 <- seq(0,lags,5) # defines position of tick marks.
  v1[1]=1 # sets the first tick mark to 1
  
  matplot(1:lags,saveeigpcor_i,type="n",xlab="lag",ylab="Absolute Value of the Eigenvalues",axes=FALSE)
  axis(1,at = v1,tck=-.05)
  axis(2,las=2)
  minor.tick(nx = 5,ny = 1,tick.ratio=0.75)
  
  grid(col = "lightgray", lty = "dotted") # To get grid behind points('type='n')
  par(new=TRUE)
  matplot(1:lags,saveeigpcor_i,type="b",xaxt="n",yaxt="n", ann=FALSE)
  abline(h = max(max_eig_ind_lag), lty="dashed", lwd=2)
  
} #closes function


####################################################### FUNCTION
################################################################
###################################################
### EIGENVECTORS OF LAGGED PARTIAL              ###
### AUTOCORRELATION MATRICES                    ###
###################################################
### INPUTS:                                     ###
### X = Data Matrix                             ###
### lags = Number of lags to test               ###
###                                             ###
### OUTPUT:                                     ###
###   ?
###                                             ###
###################################################
eigvec.p.corr <- function(X,lags){
  
  p = dim(X)[2]
  savelagcor <- array(0, dim = c(lags,p,p))
  saveeigpcor_i <- array(0, dim = c(lags,p))
  saveeigvecpcor_i <- array(0, dim = c(lags,p,p))
  
  for (i in 1:lags){
    Xlag = Xlag(X,i) #Creates a lagged matrix of data
    pcor_lag = pcor(Xlag)$estimate
    
    pcormat = matrix(rep(0),p,p) #Starting a new p.corr.matrix
    for (j in 1:p){
      for (k in 1:p){
        pcormat[j,k] =  pcor_lag[j,((i*p)+k)] # j = row, k = column
      }   
    } 
    savelagcor[i,,] = pcormat
    saveeigpcor_i[i,] = Mod(eigen(pcormat)$values)
    saveeigvecpcor_i[i,,] = eigen(pcormat)$vectors
  }
  
  par(mfrow=c(1,1))
  x_label = seq(1,lags,1)
  matplot(x_label,saveeigpcor_i,type="b",xlab="lag",ylab="Mod(Eigenvalues)",
          main="Modulus of Eigenvalues of Partial autocorrelation Matrices")
  
  saveeigvecpcor_i
} #closes function



```


In the following code snippet we load our data and difference the data as described in the main report.

```{r}
library(R.matlab)

# Loading the data from a matfile that contains all of the data sets
Data_XY <- readMat("/Users/Sebastian/Documents/Leif/DTU/Speciale/Runs/Data_XY.mat")


# Loading the data from the anaerobic digester influent and effluent and differencing it as described in the main report with respect to the prior days observation and the prior observation
AD_influent_diff <- diff(diff(scale(Data_XY$AD.influent), lag = 96), lag=1)
AD_effluent_diff <- diff(diff(scale(Data_XY$AD.effluent), lag = 96), lag=1)



# Loading the data from the different locations in the activated sludge process and differencing it as described in the main report with respect to the prior days observation and the prior observation

# For locations a-f kLa is removed since the variable value is zero, and this prevents some of the code from running, since it is unable to calculate eigenvalues
location_a_diff <- diff(diff(scale(Data_XY$location.a[,c(1,2,3,4,5,6,7,9,10,11,12,13)]), lag = 96), lag = 1)
location_b_diff <- diff(diff(scale(Data_XY$location.b[,c(1,2,3,4,5,6,7,9,10,11,12,13)]), lag = 96), lag = 1)
location_c_diff <- diff(diff(scale(Data_XY$location.c[,c(1,2,3,4,5,6,7,9,10,11,12,13)]), lag = 96), lag = 1)
location_d_diff <- diff(diff(scale(Data_XY$location.d[,c(1,2,3,4,5,6,7,9,10,11,12,13)]), lag = 96), lag = 1)
location_e_diff <- diff(diff(scale(Data_XY$location.e[,c(1,2,3,4,5,6,7,9,10,11,12,13)]), lag = 96), lag = 1)
location_f_diff <- diff(diff(scale(Data_XY$location.f[,c(1,2,3,4,5,6,7,9,10,11,12,13)]), lag = 96), lag = 1)
location_g_diff <- diff(diff(scale(Data_XY$location.g), lag = 96), lag = 1)
location_h_diff <- diff(diff(scale(Data_XY$location.h), lag = 96), lag = 1)
location_i_diff <- diff(diff(scale(Data_XY$location.i), lag = 96), lag = 1)


```


For both of the methods all lags upto lag 20 are investigated for the two methods.
The 10 in eig.p.corr(AD_influent_diff, 20, 10) corresponds to how many of the eigenvalues are shown which is equivalent to the number of variables in the dataset (both X and y), so for the AD it is 10, for the AS loc a-f it is 12 and for the AS g-i it is 13

```{r}
par(mfrow=c(1,2))
eig.p.corr(AD_influent_diff, 20, 10)
Rato.Reis.2013(AD_influent_diff,20)
```





```{r}
par(mfrow=c(1,2))
eig.p.corr(AD_effluent_diff, 20, 10)
Rato.Reis.2013(AD_effluent_diff,20)
```



```{r}
par(mfrow=c(1,2))
eig.p.corr(location_a_diff, 20, 12)
Rato.Reis.2013(location_a_diff,20)
```


```{r}
par(mfrow=c(1,2))
eig.p.corr(location_b_diff, 20, 12)
Rato.Reis.2013(location_b_diff,20)
```



```{r}
par(mfrow=c(1,2))
eig.p.corr(location_c_diff, 20, 12)
Rato.Reis.2013(location_c_diff,20)
```


```{r}
par(mfrow=c(1,2))
eig.p.corr(location_d_diff, 20, 12)
Rato.Reis.2013(location_d_diff,20)
```




```{r}
par(mfrow=c(1,2))
eig.p.corr(location_e_diff, 20, 12)
Rato.Reis.2013(location_e_diff,20)
```


```{r}
par(mfrow=c(1,2))
eig.p.corr(location_f_diff, 20, 12)
Rato.Reis.2013(location_f_diff,20)
```



```{r}
par(mfrow=c(1,2))
eig.p.corr(location_g_diff, 20, 13)
Rato.Reis.2013(location_g_diff,20)
```


```{r}
par(mfrow=c(1,2))
eig.p.corr(location_h_diff, 20, 13)
Rato.Reis.2013(location_h_diff,20)
```




```{r}
par(mfrow=c(1,2))
eig.p.corr(location_i_diff, 20, 13)
Rato.Reis.2013(location_i_diff,20)
```


